---
title: "Math 423 assignment 2"
author: "Caiwei Xiong"
date: "23/10/2020"
output: html_document
---

#### Question 3
1. 
```{r code1}
X1 = c(4,3,5,7)
Y = as.matrix(c(6,2,4,11))
fit <- lm(Y ~ X1) 
summary(fit)

```
2. 
```{r pressure}
n<- length(X1) 
X= as.matrix(cbind(rep(1,n),X1))
XtX <- t(X)%*%X
XtX_inv = solve(XtX) 
beta = XtX_inv%*% t(X)%*%Y 
beta  
```

#### Question 4
```{r is hard }
file1<-"http://www.math.mcgill.ca/yyang/regression/data/salary.csv"
salary <-read.csv(file1 ,header=TRUE)
x1<-salary$SPENDING/1000
y<-salary$SALARY
fit.Salary<-lm(y ~ x1);summary(fit.Salary)
plot(x1,y)
```
1. Verify the Estimate column
```{r is too hard}
X <- cbind(1,x1)
XTX  <- t(X)%*%X
XTy <- t(X)%*% y
beta.hat <- solve(XTX, t(X)%*%y)
beta1 <- beta.hat[2]
print(beta.hat)
```
This result as same as the output from the summary function.

2. Verify numerically the orthogonality result 
```{r is so so hard}
RP.residuals <- residuals(fit.Salary)
RP.fitted <- fitted(fit.Salary)
sum(RP.residuals)
sum(RP.residuals*(x1 - mean(x1)))
sum(RP.fitted*RP.residuals)
```

3. Compute the value of the entry  in the Std.Error column
```{r is so so so hard }
fitted.y <- X %*% beta.hat
residual.vec <- y - fitted.y
SS_Res <-  sum(residual.vec^2)
MS_Res <- SS_Res/(length(y)-2)
sigma.hat <- sqrt (MS_Res)
(estimated.covariance <- sigma.hat^2 * solve(XTX))
```
```{r is so so so hard QAQ}
(ese.vals<- sqrt(diag(estimated.covariance)))

```
This result as same as the output from the summary function. 

4. compute the Residual standard
```{r is so so so so hard}
fitted.y <- X %*% beta.hat
residual.vec <- y - fitted.y
SS_Res <-  sum(residual.vec^2)
MS_Res <- SS_Res/(length(y)-2)
sigma.hat <- sqrt (MS_Res)
sigma.hat
SS_Res
```
This result as same as the output from the summary function. 

5. Test linear association between SALARY and SPENDING
```{r is so so so so so hard}
coef(fit.Salary)
cor.test(x1,y,mehtod=c("person", "kendall", "spearman"))

```


6.
```{r does not know how to write}
confint(fit.Salary, level = 0.99)
tstat = (beta1)/ 311.7043
tstat
mquantile = qt(1-0.01/2, df = length(y)-2)
p_value = 2*pt(-abs(tstat), df =  length(y)-2)
p_value
```

7. F- statistic 
```{r is so so so so so so hard} 
SS_T <- (length(y)-1 ) * var(y)
SS_R <- SS_T - SS_Res 
SS_T
SS_R
(R_squared <- SS_R / SS_T) 
Fstat <- (SS_R/ (2-1)) / (SS_Res/(length(y)-2))
Fstat 
anova(fit.Salary)

summary(fit.Salary)$fstatistic
pf(Fstat, 1, length(y)-2, lower.tail = FALSE)
```

8. S_xy
```{r wow} 
S_xy <- sum((y - mean(y))* (x1 - mean(x1)))
S_xy
```

9. spending per pupil is 5000
```{r is fantastic} 
y <- 12129.371 + 3307.585*5.0
y
```


10.
```{r kills me} 
(x1new <- matrix(c(1,5.0), nrow = 1))
pred.var <- sigma.hat^2 * x1new %*% solve(XTX) %*% t(x1new) 
pred.var
(sigma.hat_1<-sqrt(pred.var))
```

11.
```{r kills me me }
k = qt(1-0.025, df= length(x1)-2 )
CIlower <- sum(summary(fit.Salary)$coef[,1]*c(1,5))- k* (sigma.hat_1) 
CIupper <- sum(summary(fit.Salary)$coef[,1]*c(1,5))+ k* (sigma.hat_1)
c("CIlower" = CIlower, "CIUpper" = CIupper)

sigma_1 <- sigma.hat^2*(1+(1/length(x1)) + (((5-mean(x1))^2)/(var(x1)*(length(x1)-1))))
PIlower <- sum(summary(fit.Salary)$coef[,1]*c(1,5)) - k*(sigma_1^(1/2))
PIupper <- sum(summary(fit.Salary)$coef[,1]*c(1,5)) + k*(sigma_1^(1/2)) 

c("PIlower" = PIlower, "PIUpper" = PIupper)                              

predict(fit.Salary, newdata = data.frame(x1=5.0), interval = "confidence")
predict(fit.Salary, newdata = data.frame(x1=5.0), interval = "prediction") 

```


#### Question 5
1.
```{r part 1}
data.source <- "http://www.math.mcgill.ca/yyang/regression/data/genetic_data.csv"
genetic_data <- read.csv(file = data.source)
head(genetic_data)
dim(genetic_data)
x<- genetic_data[,2]
y<- genetic_data[,1]
m_original <- lm(y~x)
plot(y~x, xlab = "Gene_x", ylab = "Gene_y", main="Original data")
abline(m_original)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
#print(tstat_vec)
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
hist(tstat_vec, main="Histgram of permuted t-statistic", breaks = 100, xlim = c(-20,20))
abline(v= - tstat_original, col="red", lwd = 4)
abline(v= tstat_original, col="red", lwd = 4)
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

2.
```{r part 2}
# this method does not work 
y<- genetic_data[,1] 
permutation_test_2 <- function(output, input, nrep  = 1000){
  for (i in 3:11){
    tstat_vec <- rep(0, nrep)
    x <-genetic_data[,i]
    #print(i)
    for(j in 1:nrep){
      y_permuted <- sample(output) 
      m_permuted <- lm(y_permuted ~ input) 
      results <- summary(m_permuted)
      tstat_vec[j] <- results$coef[2,3]
      #print(i*j)
      #print(tstat_vec[j])
    }
  }
#  print(tstat_vec)
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test_2(y,x,nrep = M) 
par(mfrow = c(1,1))
hist(tstat_vec, main="Histgram of permuted t-statistic", breaks = 100, xlim = c(-20,20))
abline(v= - tstat_original, col="red", lwd = 4)
abline(v= tstat_original, col="red", lwd = 4)
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for gtwo}
data.source <- "http://www.math.mcgill.ca/yyang/regression/data/genetic_data.csv"
genetic_data <- read.csv(file = data.source)
x<- genetic_data[,3]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p

```
```{r for gthree }
x<- genetic_data[,4]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```
```{r for gfour} 
x<- genetic_data[,5]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p

```

```{r for gfive}
x<- genetic_data[,6]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for gsix}
x<- genetic_data[,7]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for gseven}
x<- genetic_data[,8]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for geight}
x<- genetic_data[,9]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for gnine}
x<- genetic_data[,10]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r for gten}
x<- genetic_data[,11]
y<- genetic_data[,1]
m_original <- lm(y~x)
results<- summary(m_original)
results$coef
(tstat_original <- results$coef[2,3])

permutation_test <- function(output, input, nrep  = 1000){
  tstat_vec <- rep(0, nrep)
  for(i in 1:nrep){
    y_permuted <- sample(output) 
    m_permuted <- lm(y_permuted ~ input) 
    results <- summary(m_permuted)
    tstat_vec[i] <- results$coef[2,3]
  }
  return(tstat_vec)
}
M = 1000
tstat_vec<- permutation_test(y,x,nrep = M) 
par(mfrow = c(1,1))
p <- sum(tstat_vec > abs(tstat_original)| tstat_vec < - abs(tstat_original))/M
p
```

```{r part4} 
summary(lm(genetic_data[,2]  ~ genetic_data[,1]))
summary(lm(genetic_data[,3]  ~ genetic_data[,1]))
summary(lm(genetic_data[,4]  ~ genetic_data[,1]))
summary(lm(genetic_data[,5]  ~ genetic_data[,1]))
summary(lm(genetic_data[,6]  ~ genetic_data[,1]))
summary(lm(genetic_data[,7]  ~ genetic_data[,1]))
summary(lm(genetic_data[,8]  ~ genetic_data[,1]))
summary(lm(genetic_data[,9]  ~ genetic_data[,1]))
summary(lm(genetic_data[,10]  ~ genetic_data[,1]))
summary(lm(genetic_data[,11]  ~ genetic_data[,1]))
```

